---
title: "Assignment 1 Part 2"
author: "Krista Bennatti-Roberts"
date: "October 6, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

2. This question explores how Magna sales (millions of US dollars) are affected by US GDP (billions of US dollars). Use 2000:1 to 2014:4 for training and 2015:1 to 2017:1 for testing. The data are in as_2.csv. 

```{r message=FALSE}
library(fpp)
library(knitr)
magna = read.csv('C:/schulich/Econ R/Assignments/as_2.csv')
```

View the data and split into training and testing
```{r}
head(magna)
tail(magna)
# tell R this is a time series
mgts = ts(magna, start = c(2000,1), end = c(2017,1), frequency = 4)
# split into training and testing
mtraints = window(mgts, start = c(2000, 1), end = c(2014, 4)) 
mtestts = window(mgts, start = c(2015, 1), end = c(2017, 1))
```

Plot the data for fun
```{r}
plot(MGA ~ GDP, data = mtraints, xlab = "GDP", ylab = "MGA Sales", main = "Sales vs. GDP")
```


a) Using the training data to estimate a linear regression with Magna sales as the dependent variable and GDP as the explanatory variable. How does this regression fit? 8 points

```{r}
#linear model, sales as dependent variable
fit = tslm(MGA ~ GDP, data = mtraints)
plot(MGA ~ GDP, data = mtraints, xlab = "GDP", ylab = "MGA Sales", main = "Linear Model")
abline(fit)
```

Comment: Visually, line fits data reasonably well. Some outliers around GDP = 14000.

Explore fit of model some more
```{r}
summary(fit)
anova(fit)
```

Comments:
Adjusted R squared shows that GDP explains approximately 86% of the variability in sales, implying a relatively good fit.



Plot residuals
```{r}
res = residuals(fit)
plot(res, type = "p", xlab = "Years", ylab = "Residuals", main = "Residual Plot")
abline(0,0)
```



Comment: The residuals have no prominent trend and appear reasonably random, implying an okay fit. However there may be a slight cylical trend present and there are more positive residuals than negative. There are also a few notable outliers.

More Residual Plots
```{r}
par(mfrow = c(2,2))
plot(fit)
par(mfrow = c(1,1))
```

Comment: Q-Q plot shows resdiduals tend to cluster around line (indicating normalility) with the exception of a number of notable outliers. 
There are three outliers identified (#37, 38, 39) by scale-location, residuals vs fitted and the Q-Q plots,
Two influential outliers identified by Cook's distance (#36, 54).

b) Use scenario forecasting to forecast Magna sales over the test period. Use a random walk with drift to forecast GDP over the test period. 4 points

Create GDP forecast using RWF with drift for scenario forecast
```{r}
h = length(mtestts[,'GDP'])
h
fcastgdp = rwf(mtraints[,'GDP'], h=h, drift = TRUE)
print(fcastgdp)
plot(fcastgdp, xlab = "Year", ylab = 'GDP', main = 'GDP Forecast using RW w Drift')
```

Forecast sales using GDP forecast as input
```{r}
fcastsales = forecast(fit, newdata = data.frame(GDP = c(fcastgdp$mean)))
print(fcastsales)
plot(fcastsales, xlab = "Year", ylab = 'MGA Sales', main = "Scenario Forecast Based on GDP RWF w/ Drift")
```

c) Use a linear trend to forecast Magna sales for the test period. 4 points

```{r}
salesfit = tslm(MGA ~ trend, data = mtraints)
fcastlin = forecast(salesfit, h=h)
print(fcastlin)
plot(fcastlin, xlab = "Year", ylab = 'MGA Sales', main = "Linear Trend Forecast of Magna Sales" )
```

d) Compare the accuracy of the two forecasting models. Which model forecasts the best? 4 points

```{r}
#Accuracy of linear model
acclin = print(accuracy(fcastlin$mean, mtestts[,'MGA']))
#Accuracy of scenario model
accsen = print(accuracy(fcastsales$mean, mtestts[,'MGA'] ))
```
Comment: The linear trend forecast is the better forecaset based on the RMSE metric 
