---
title: "Assignment 3 Parts 2 and 3"
author: "Krista Bennatti-Roberts"
date: "October 25, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

2. This question uses the Staples sales data from Assignment 1. Use the same training and test data sets that you used for Assignment 1. 

For this data set determine the best forecasting method (average, naïve, seasonal naïve, linear trend, STL decomposition, exponential smoothing, Holt-Winters' method, ETS). Using the best model, forecast Staples sales for 2017:3 to 2018:2. 10 points

Get necessary libraries:

```{r message=FALSE} 
library(fpp)
library(knitr)
```

Set up training and testing:

```{r}
stp_hd = read.csv("C:/schulich/Econ R/Assignments/as1_data.csv")
series = ts(stp_hd, start = 2000, frequency = 4)
train = window(series, start = c(2000,1), end = c(2013, 4))
test = window(series, start = c(2014, 1), end = c(2017, 2))
head(series)
```


STAPLES


Set up
```{r}
trainstp = train[,4]
teststp = test[,4]
tail(trainstp)
tail(teststp)
h = length(teststp)
h
```

Run the models:

```{r}
#average, naive, seasonal naive
fitmean = meanf(trainstp, h=h)
fitnaive = naive(trainstp, h=h)
fitsnaive = snaive(trainstp, h=h)
#linear trend
tslmstp = tslm(trainstp ~ trend)
fittslm = forecast(tslmstp, h=h)
#stl
stlstp= stl(trainstp, t.window=15, s.window="periodic", robust=TRUE)
fitstl= forecast(stlstp, method='rwdrift', h=h)
#exponential
fitses = ses(trainstp, h=h)
#holt-winter's multiplicative
fithw_m = hw(trainstp, seasonal="multiplicative", h=h)
#holt-winter's damped
fithw_m_d = hw(trainstp, seasonal="multiplicative", damped = TRUE, h=h)
#error trend seasonl (ETS)
etsstp = ets(trainstp, model = "ZZZ")
fitets = forecast(etsstp, h=h)
```

Determine accuracy

```{r}
am = accuracy(fitmean, teststp)
an= accuracy(fitnaive, teststp)
asn= accuracy(fitsnaive, teststp)
atslm= accuracy(fittslm, teststp)
astl = accuracy(fitstl, teststp)
ases= accuracy(fitses, teststp)
ahwm= accuracy(fithw_m, teststp)
ahwmd= accuracy(fithw_m_d, teststp)
aets= accuracy(fitets, teststp)

#combine accuracy into table

accTable = rbind(am,an,asn,atslm, astl, ases,ahwm, ahwmd,aets)
row.names(accTable)= c("Mean Train", "Mean Test","Naive Train", "Naive Test", 
                       "SNaive Train", "SNaive Test", "Linear Train", "Linear Test", "STL Train", "STL Test",
                       "SES Train", "SES Test","HW Mult Train", "HW Mult Test", 
                       "HW Damped Train", "HW Damped Test", "ETS Train", "ETS Test")
#order by MASE
accFrame = as.data.frame(accTable)
print(accFrame[order(accFrame$MASE),])
```

Holt-Winter's (multiplicative) is the best model according to the MASE metric

Forecast sales for 2017 Q3 through 2018 Q2 using the best model

```{r}
hw_m_future= hw(series[,4], seasonal="multiplicative", h=4)
summary(hw_m_future)
```

HOME DEPOT

Set up:
```{r}
trainhd = train[,3]
testhd = test[,3]
tail(testhd)
h = length(teststp)
h
```


Run the models

```{r}
#average, naive, seasonal naive
fitmean = meanf(trainhd, h=h)
fitnaive = naive(trainhd, h=h)
fitsnaive = snaive(trainhd, h=h)
#linear trend
tslmstp = tslm(trainhd ~ trend)
fittslm = forecast(tslmstp, h=h)
#stl
stlstp= stl(trainhd, t.window=15, s.window="periodic", robust=TRUE)
fitstl= forecast(stlstp, method= 'rwdrift', h=h)
#exponential
fitses = ses(trainhd, h=h)
#holt-winter's multiplicative
fithw_m = hw(trainhd, seasonal="multiplicative", h=h)
#holt-winter's damped
fithw_m_d = hw(trainhd, seasonal="multiplicative", damped = TRUE, h=h)
#error trend seasonl (ETS)
etslbr = ets(trainhd, model = "ZZZ")
fitets = forecast(etslbr, h=h)
```

Determine accuracy of the models

```{r}
am = accuracy(fitmean, testhd)
an= accuracy(fitnaive, testhd)
asn= accuracy(fitsnaive, testhd)
atslm= accuracy(fittslm, testhd)
astl = accuracy(fitstl, testhd)
ases= accuracy(fitses, testhd)
ahwm= accuracy(fithw_m, testhd)
ahwmd= accuracy(fithw_m_d, testhd)
aets= accuracy(fitets, testhd)

#combine accuracy into table

accTable = rbind(am,an,asn,atslm, astl, ases,ahwm, ahwmd,aets)
row.names(accTable)= c("Mean Train", "Mean Test","Naive Train", "Naive Test", 
                       "SNaive Train", "SNaive Test", "Linear Train", "Linear Test", "STL Train", "STL Test",
                       "SES Train", "SES Test","HW Mult Train", "HW Mult Test", 
                       "HW Damped Train", "HW Damped Test", "ETS Train", "ETS Test")
#order by MASE
accFrame = as.data.frame(accTable)
print(accFrame[order(accFrame$MASE),])
```

STL is the best (most accurate) model according to the MASE metric

Forecast sales for 2017 Q3 through 2018 Q2 
```{r}
stlstp= stl(series[,3], t.window=15, s.window="periodic", robust=TRUE)
fitstl= forecast(stlstp, method= 'rwdrift', h=4)
summary(fitstl)
```

